Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/WSSSPE3_review.tex
small changes, line breaks

The workshop included multiple mechanisms for participation and encouraged team
building around solutions. It strongly encouraged participation of early-career
scientists, postdoctoral researchers, and graduate students, with funds provided
to the conference organizers by the Moore Foundation, the National Science
Foundation (NSF), and the Software Sustainability Institute, to support the
travel of potential participants who would not otherwise be able to attend,
younger participants, and those from underrepresented groups. These funds allowed
16 additional participants to attend.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{need to update for WSSSPE3}
\todo{KEN: Should these be bolded? - dsk: yes, titles can be bolded, but contents under each should not}
\begin{comment}
\note{
\href{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/}{Slides.}}
\end{comment}

\begin{enumerate}
\item \textbf{Benjamin Tovar and Douglas Thain: \textit{Freedom vs. Stability:
Facilitating Research Training While Supporting Scientific Research}}
A case study: ``The Cooperative Computing Lab'' at the University of Notre Dame
is a small group of individuals with the main tasks of collaborating with
people that have large-scale computing problems, operating various parallel
computer systems, conducting computer science research, and developing open
source software. One of the main challenges they face is finding the balance
between flexibility/training and stability/quality. Their current solution for
ensuring the latter was to add a software engineer (the presenter) to the
existing team of faculty and students, who now also serves as ``spring''
between flexibility and stability.

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of
computing, which is evidenced by the increase in the number of events that
focus on the topic - nevertheless, it isn’t well understood yet. Penzenstadler
argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a
range of different dimensions including environmental, economic, individual,
social and technical. She introduced the Karlskrona Manifesto on Software
Design, which distills knowledge from a broad range of related work on the
topic of sustainability into a set of (mis-)perceptions and principles. The
manifesto does not proclaim that there is an easy, one size fits all solution
around the corner, but rather points out that sustainability is a ``wicked
problem'' and is often misunderstood. Due to these misperceptions, even though
sustainability's importance is increasingly recognized, many software systems
are unsustainable. Even more alarming is that most software systems' broader
impacts on sustainability are unknown. To change this, the Karlskrona Manifesto
proposes nine principles and commitments. These commitments are not dogmatic
laws, but rather a commitment to rethink, to move beyond the silo mentality,
and analyze in more depth. As such, they do not restrict, but rather open up a
space for discussion.

@@

\item \textbf{Jeffrey Carver: \textit{On the Need for Software Engineering Support for
Sustainable Scientic Software}}
Jeff argued that for scientific software to be truly sustainable, there is a need
for developers to use appropriate software engineering practices. His experience
interacting with scientific teams indicates that choosing and tailoring these
practices is not a trivial exercise. There is a general culture clash between
software engineering and science that hinders our ability to communicate and
choose appropriate methods. In addition, many experience scientific software
developers appear to be unaware of software engineering practices that may be
beneficial to them. The most appropriate software engineering practices are
those that are lightweight, properly tailored, and focus on the key software
development problems faced by scientists. In order to increase the use of
software engineering in science, we need more documented success stories. These
successes need to be socialized within the scientific community through
workshops like the Software Engineering for Science workshop
series\footnote{\href{http://www.SE4Science.org/workshops}{http://www.SE4Science.org/workshops}}
and the new Software Engineering track in
\textit{Computing in Science and Engineering} magazine.

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
%\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
%\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
%\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}

>>>>

The workshop included multiple mechanisms for participation and encouraged team
building around solutions. It strongly encouraged participation of early-career
scientists, postdoctoral researchers, and graduate students, with funds provided
to the conference organizers by the Moore Foundation, the National Science
Foundation (NSF), and the Software Sustainability Institute, to support the
travel of potential participants who would not otherwise be able to attend,
early-stage researchers, and those from underrepresented groups. These funds allowed
16 additional participants to attend.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{need to update for WSSSPE3}
\todo{KEN: Should these be bolded?}
\begin{comment}
\note{
\href{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/}{Slides.}}
\end{comment}

\begin{enumerate}
\item \textbf{Benjamin Tovar and Douglas Thain: \textit{Freedom vs. Stability:
Facilitating Research Training While Supporting Scientific Research}}

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}

@@

\item \textbf{Jeffrey Carver: \textit{On the Need for Software Engineering Support for
Sustainable Scientic Software}}

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
%\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
%\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}
